### 2019-09-19 initial/notes.tex
Finalized plotting for now
0919 draft meeting w/ dong, wrote out adiabiticity constraint

@@

We can investigate the values of $\theta_f$ given some initial mutual
inclination $\hat{s} \cdot \hat{l}_p \equiv \theta_{sp, i}$ ($J_i = 2\pi\p*{1 -
\cos \epsilon}$ where $\epsilon$ is the initial mutual inclination). To do this,
we need to understand what the possible transitions are. When the separatrix
first appears to the final state, the transitions are $A_2 \to A_3, A_2 \to A_1,
A_3 \to A_1, A_3 \to A_2 \to A_1, A_3 \to A_3$. Let's reproduce the analytical
forms of these three areas below:
\begin{subequations}\label{se:}
    \begin{align*}
        z_0 &= \eta\cos I &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1} &
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1},
    \end{align*}
    \begin{align}
        A_2 &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_1 &= 2\pi\p*{1 - z_0} - \frac{A_2}{2},\\
        A_3 &= 2\pi\p*{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
To determine which of the transition histories we fall under (to get an
analytical prediction for $\cos \theta_f$), we need to determine $\eta_\star$,
so we know how the enclosed $J$ by the trajectory changes due to separatrix
encounter. A guess for numerical root-finding can be put together for
trajectories that encounter from $A_2$ via (inverting the separatrix area
formula for small $\eta$)
\begin{equation*}
    \eta_\star\p*{A_2} = \p*{A_2/16}^2 / \sin I,
\end{equation*}
while for trajectories that encounter from $A_3$ we can solve the quadratic $A_3
= 2\pi \eta \cos I - 8\sqrt{\eta \sin I} = 4\pi - J$ and obtain
\begin{equation*}
    \sqrt{\eta_\star} = \frac{4\sqrt{\sin I} + \sqrt{16\sin I + 2\pi \cos I
        A_3}}{ 2\pi \cos I}.
\end{equation*}
This will often run over $\eta_c$, so we guess the smaller of $\eta_\star,
\eta_c$.

>>>>

\section{Non-adiabatic Regime}

\subsection{Decaying Disk: Dong Notes}

In the non-adiabatic regime, the Hamiltonian is much less useful, so we return
to EOM in corotating frame (after dividing through by $\Omega_{sl}$ compared to
Dong's notes)
\begin{equation}
    \rd{\hat{s}}{t} = \p*{\hat{s} \cdot \hat{l}}\p*{\hat{s} \times \hat{l}}
        + \eta \cos I \p*{\hat{l}_d \times \hat{s}}.
\end{equation}
We adopt convention where $\Omega_{sl}, \Omega_{lp} > 0, \eta =
\frac{\Omega_{ld}\cos I}{\Omega_{sl}}$, and time units are really $t \equiv
\Omega_{sl}t$. In this decaying disk problem, $\eta\p*{-\infty} \gg 1,
\eta\p*{+\infty} \ll 1$.

Define again $\hat{l} = \hat{z}, \hat{l}_d = \hat{z} \cos I + \hat{x}\sin I$,
then we go immediately into coordinates
\begin{equation}
    \rd{\hat{s}}{t} = \s*{
        \p*{\eta \cos I - 1}\hat{z} + \eta \sin I \hat{x}} \times \hat{s}.
\end{equation}

Now, let's assume that $s_z \approx 1$ throughout the evolution of $\hat{s}$
(maybe this can be relaxed later), which also requires $I \approx 0$. Then let's
examine the evolution of quantity $S = s_x + is_y$ instead:
\begin{equation}
    \rd{S}{t} = i\p*{\eta\cos I - 1}S - i \eta \sin I.\label{eq:nonad_ode}
\end{equation}
Now this is a first-order ODE in $S$, albeit complex, which can be solved via
an integrating factor
\begin{align}
    \Phi(t) &\equiv \int_{-\infty}^t \p*{1 - \eta(t') \cos I}
        \mathrm{d}t',\\
    \at{S(t) e^{i\Phi(t)}}_{-\infty}^t
        &= \int\limits_{-\infty}^t e^{i\Phi(t')}
            \p*{-i\eta(t')\sin I}\;\mathrm{d}t'\label{eq:nonad_int}
\end{align}
We now invoke stationary phase, asserting that $e^{i\Phi(t)}$ is dominated by
its contribution where $\dot{\Phi} = 0$ (the phases add constructively). But
$\dot{\Phi} = 0$ is where $1 - \eta\cos I = 0$, or where $\eta\cos I = 1$.

Now at this point, let's choose $\eta(0) = 1/\cos I, \at{\rd{\eta}{t}}_{t=0} =
-\epsilon/\cos I$. Then we expand near $t = 0$ so
\begin{align*}
    \Phi(t) &\approx \Phi(0) + \frac{1}{2}\ddot{\Phi}t^2,\\
        &\approx \Phi(0) + \frac{1}{2}\epsilon t^2,\\
    \int\limits_{-\infty}^t e^{i\Phi(t')}\eta(t')\;\mathrm{d}t'
        &\approx
        \begin{cases}
            0 & t < 0,\\
            \frac{1}{\cos I}e^{i\Phi(0)}\int\limits_{-\infty}^\infty
                \exp\s*{\frac{i}{2}\epsilon t^2}\;\mathrm{d}t
                & t > 0.
        \end{cases}\\
    \int\limits_{-\infty}^\infty
                \exp\s*{\frac{i}{2}\epsilon t^2}\;\mathrm{d}t
        &= \int\limits_{-\infty}^\infty e^{-\tau^2}\;\mathrm{d}\tau
            \sqrt{\frac{2}{i\epsilon}},\\
        &= \sqrt{\frac{2\pi}{i\epsilon}}.
\end{align*}

Now, it should be noted that $e^{i\Phi}$ is just a phase; all we really care
about is $\abs*{S} = \sqrt{1 - s_z^2}$. Thus, taking the absolute value of both
sides of \autoref{eq:nonad_int} we obtain
\begin{align}
    \abs*{S}(t) &= \tan I\sqrt{\frac{2\pi}{\epsilon}}\Theta(t).
        \label{eq:nonad_dong}
\end{align}
It should be noted that $\theta_{sl, f} \approx \abs*{S}(\infty)$, and so we
recover Dong's result $\theta_f = \sqrt{2\pi/\epsilon}\tan I$.

To compare to data, let's define a few more things: $\theta_{sl, f} = \hat{s}
\cdot \hat{l}(t \to \infty)$ and $\theta_{sd, i} = \hat{s} \cdot \hat{l}_d(t \to
-\infty)$ (note that $I \equiv \theta_{ld}$ is constant). In our derivation
above, we've sort of assumed $I \approx 0$, so $\theta_{sd, i} \approx
\theta_{sl, i} = 0$ is the effective initial condition above.

Then, for some $\theta_{sd, i} > 0$, we can intuit as follows: $\hat{s}$
initially librates around $\hat{l}_d$ with oscillation amplitude $\theta_{sd,
i}$. So if $\theta_{sd, i} = 0$ produces \autoref{eq:nonad_dong}, we might
expect small deviations of amplitude $\theta_{sd, i}$ about CS2 to be ``frozen
in'' as the disk dissipates non-adiabatically. Thus, $\theta_{sl, f} \in
\theta_{sl, f, 0} + \s*{\pm \theta_{sd, i}}$.

More importantly though, we want to find leading order corrections to the above
result for: nonzero $I$, and somewhat smaller $\epsilon$. The problems with
Dong's result can be seen in the two figures in \autoref{fig:nonad_scan}, which
are derived for $\theta_{sd, i} \approx 0$. In this limit, $\theta_{sl, i} = I$,
which explains why the predictions break down somewhat when $I$ is no longer
small. Furthermore, for smaller $\epsilon$, the stationary phase approximation
breaks down. We investigate the effect of these below.
\begin{figure}[t]
    \centering
    \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{2_toy2/3scan.png}
    \end{subfigure}

    \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{2_toy2/3scan_20.png}
    \end{subfigure}
    \caption{$I = 5^\circ, 20^\circ$ distributions vs data. Red lines are the
    result of Dong's calculation, blue are with my
    improvements.}\label{fig:nonad_scan}
\end{figure}

\subsection{Extending Dong's result}

Perhaps the easiest correction to make is, instead of assuming $s_z = 1$, to
hold it constant at $s_z \approx \cos I$, which just changes the inhomogeneous
term in \autoref{eq:nonad_ode} to $\p*{-i\eta \sin I \cos I}$; this is a very
small change since $\cos I \approx 1$ for nearly all physical parameter space,
but does improve the fit slightly.

One other observation is that the LHS of \autoref{eq:nonad_int} should really be
$\abs*{S\p*{t}e^{i\Phi(t)} - S\p*{-\infty}e^{i\Phi(-\infty)}}$, but we don't
really know $e^{i\Phi}$ since it oscillates so quickly. Thus, only when $S\p*{t}
\gg S\p*{-\infty}$ or $\theta_{sl, f} \gg I$ is the LHS easily evaluated. But at
the very least we can claim that $\theta_{sl, f} = \max\p*{\abs*{S}, I}$, so
that at very small $\abs*{S}$ we just use $I$ the approximate location of CS2.
This seems to be in pretty good agreement if we look at
\lstinline{2_toy2/3Iscan} plots.

Somewhat surprisingly, using a proper $\theta_{sl, f} = \arccos 1 - S^2$
actually hurts the fit; this would manifest in the larger $\theta_{sl, f}$
regime, so it appears increasing adiabiticity dominates this correction.

The other correction to consider is how the nonadiabatic assumption breaks down.
In the stationary phase integral, we assumed that the phase could be integrated
as an imaginary Gaussian.

\textbf{Take 1 (fail):} This is only true if the in the
vicinity of the minimum, $\Phi(t)$ is approximately symmetric. The ``vicinity''
here is defined as where $\Phi(t) \lesssim \Phi_{\min} + 2\pi$, since that is
where the phase begins to oscillate quickly again. To simplify, we can write
$\eta = e^{-\epsilon t}/\cos I$, then we can integrate analytically:
\begin{equation}
    \Phi(t) = t + \frac{e^{-\epsilon t}}{\epsilon}
        \approx \frac{1}{\epsilon} + t - te^{-\epsilon t} + \mathcal{O}(t^2).
\end{equation}
can be written down analytically, we can examine how symmetric $\Phi$ is about
its minimum. It turns out though that $\Phi$ is actually \emph{more} symmetric
for smaller $\epsilon$, so stationary phase should work \emph{better}. Hm.
Similar arguments can be constructed for $\Phi$ being symmetric about $\pm
\sigma \sim 1/\sqrt{\epsilon}$, or for the variation of $\eta$ across $t \in \pm
1/\sqrt{\epsilon}$, but they all seem to favor smaller $\epsilon$ being more
accurate.

\textbf{Millholland Non-Adiabiticity:} The criterion for adiabiticity presented
in Millholland \& Batygin \emph{Primordial Obliquities} is (their Equation 15)
\begin{equation}
    \dot{\alpha} - \dot{g} \lesssim \alpha g \sin \theta \sin I.
\end{equation}
Noting $\dot{\eta} = \p*{\frac{\dot{g}}{g} - \frac{\dot{\alpha}}{\alpha}}\eta$,
where $\eta = \frac{g\cos I}{\alpha}$ ($g, \alpha$ follow their definitions),
and that at crossing they define $g \approx \alpha$, then
\begin{align*}
    \frac{\dot{\alpha}}{\alpha} - \frac{\dot{g}}{g}
        &\lesssim g\sin\theta\sin I,\\
    -\s*{\frac{\dot{\alpha}}{\alpha} - \frac{\dot{g}}{g}}\cos I = \dot{\eta}
        = -\epsilon \alpha \eta &\gtrsim -g\sin\theta\sin I\cos I,\\
    \epsilon &\lesssim \frac{\sin \theta \sin I \cos I}{\eta}
        \approx \sin\theta \sin I.
\end{align*}
This is overplotted in the vertical blue line in \autoref{fig:nonad_scan}.

@@

We can investigate the values of $\theta_f$ given some initial mutual
inclination $\hat{s} \cdot \hat{l}_p \equiv \theta_{sp, i}$ ($J_i = 2\pi\p*{1 -
\cos \epsilon}$ where $\epsilon$ is the initial mutual inclination). To do this,
we need to understand what the possible transitions are. When the separatrix
first appears to the final state, the transitions are $A_2 \to A_3, A_2 \to A_1,
A_3 \to A_1, A_3 \to A_2 \to A_1, A_3 \to A_3$. Let's reproduce the analytical
forms of these three areas below:
\begin{subequations}
    \begin{align*}
        z_0 &= \eta\cos I &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1} &
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1},
    \end{align*}
    \begin{align}
        A_2 &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_1 &= 2\pi\p*{1 - z_0} - \frac{A_2}{2},\\
        A_3 &= 2\pi\p*{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
To determine which of the transition histories we fall under (to get an
analytical prediction for $\cos \theta_f$), we need to determine $\eta_\star$,
so we know how the enclosed $J$ by the trajectory changes due to separatrix
encounter. A guess for numerical root-finding can be put together for
trajectories that encounter from $A_2$ via (inverting the separatrix area
formula for small $\eta$)
\begin{equation*}
    \eta_\star\p*{A_2} = \p*{A_2/16}^2 / \sin I,
\end{equation*}
while for trajectories that encounter from $A_3$ we can solve the quadratic $A_3
= 2\pi \eta \cos I - 8\sqrt{\eta \sin I} = 4\pi - J$ and obtain
\begin{equation*}
    \sqrt{\eta_\star} = \frac{4\sqrt{\sin I} + \sqrt{16\sin I + 2\pi \cos I
        A_3}}{ 2\pi \cos I}.
\end{equation*}
This will often run over $\eta_c$, so we guess the smaller of $\eta_\star,
\eta_c$.
